---
title: "Computing returns and sample stats"
author: "Kirsten Zimbardi"
date: "15 July 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clean workspace
rm(list=ls())

# packages
require(lubridate)
require(MASS)

# functions
source("HelperFunctionsPorfolioMgmt.R")

```

load data 
```{r}
# multiple files
#fileNames <- Sys.glob("*.csv")
#col.names = c("Date", "Amount", "Description", "Balance")
#data.ls = NULL
#for (i in 1:length(fileNames)) {
#  data.ls[[i]] <- read.csv(fileNames[i], header = F, stringsAsFactors = F)
#  names(data.ls[[i]]) = col.names
#  data.ls[[i]]$Date = as.POSIXlt(data.ls[[i]]$Date, format = "%d/%m/%Y")
#}

# single file
data <- read.csv("Data.csv", header = T, stringsAsFactors = F)

cols.price = c(2,3)
for (i in cols.price) {
  names(data)[i] = paste0(names(data)[i], ".Price")
}

data$Date = as.POSIXlt(data$Date, format = "%d/%m/%Y")


```


computing returns  
```{r}
df = data
for (j in cols.price) {
  for (i in 2:nrow(df)) {
    df[i,(j+2)] = df[i,j]/df[(i-1),j] - 1
  }
  names(df)[(j+2)] = gsub(".Price", ".Returns", names(df)[(j)])
}

data = df
```


computing average returns, variance and covariance  
```{r}
df = data
cols.return = c(4:5)

m = NULL
v = NULL
for(j in 1:length(cols.return)) {
  m[j] = mean(df[,cols.return[j]], na.rm=T)
  v[j] = var(df[,cols.return[j]], na.rm=T)
}

cv = cov(df[2:nrow(df),cols.return[1]], df[2:nrow(df), cols.return[2]])


df.stats = Stats(df, cols.return)
df.stats

df$period = as.factor(ifelse(df$Date>"2003-12-31", "2004-2015", "1992-2003"))

stats.spx.index = Stats.grouped(df, 4, 6)
stats.spx.index *100
stats.us.index = Stats.grouped(df, 5, 6)
stats.us.index *100

```

preference metric - sample  
?calculates preference-equivalent returns  
```{r}


pref.sample.percent(exp.return=2, variability=20, v.type="vol")
pref.sample.percent(2, 20, v.type="vol")

pref.sample(exp.return=df.stats[1,1], variability=df.stats[1,2], risk.aversion=10, v.type="var")
pref.sample(exp.return=df.stats[2,1], variability=df.stats[2,2], risk.aversion=10, v.type="var")

```

preference metric - probability  
```{r}
E.vec = c(1, 2, 3)/100
p.vec = c(1, 2, 3)/100
sum(E.vec * p.vec)

```

expected return - sample  
```{r}
#r = c(0.15, 0.13, -0.2, 0.12, 0.02)
Return = c(15, 13, -20, 12, 2)/100

Er = 1/length(Return) * sum(Return)
Er
```

total return  
```{r}

Price.initial = 1000
Price.later = 1000
cap.gain = 0
dividends = 50
(Price.initial - Price.later + cap.gain + dividends) / Price.later

```


total return with inflation rate  
```{r}
Price.initial = 100
Inflation.initial = 1
Price.later = 90
Inflation.later = 1 - (5/100)
dividends.later = 0

(((Price.later + dividends.later)/Inflation.later) - (Price.initial/Inflation.initial)) / (Price.initial/Inflation.initial)
```


expanding average returns, variance and covariance  
```{r}
start = which(df$Date == "1996-01-31")
a = rep(0,(nrow(df) - start + 1))
exp.ave = as.data.frame(cbind(a, a))
exp.var = as.data.frame(cbind(a, a))
exp.cov = as.data.frame(a)
for(j in 4:5) { 
  for(i in start:nrow(df)) {
    exp.ave[(i-start+1),(j-3)] = mean(df[2:i,j])
    exp.var[(i-start+1),(j-3)] = var(df[2:i,j])
  }
}

# visualising trends over time
dates = df[50:nrow(df),1]
equity = df2[,4]
bonds = df2[,5]

quartz()
plot(dates, equity, type="n")
lines(dates, equity, col="red", lwd=2)
lines(dates, bonds, col="blue", lwd=2)


equity = exp.ave[,1]
bonds = exp.ave[,2]

quartz()
plot(dates, equity, type="n")
lines(dates, equity, col="red", lwd=2)
lines(dates, bonds, col="blue", lwd=2)


equity = exp.var[,1]
bonds = exp.var[,2]*10
quartz()
plot(dates, equity, type="n")
#lines(dates, equity, col="red", lwd=2)
lines(dates, bonds, col="blue", lwd=2)

plot(dates, bonds)

exp.cov = NULL
for(i in start:nrow(df)) {
    exp.cov[(i-start+1)] = cov(df[2:i,4], df[2:i,5])
}

plot(dates, exp.cov)

  
```


# Week 2  
## Optimising portfolio construction  
Maximising Er by altering w  
```{r}
# expected return - portfolio
Er.p = Rrisk.free + t(w)


```


